---
title: "Co-authors"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Co-authors

```{r echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
library(easyPubMed)
library(tidyverse)
library(igraph)
library(visNetwork)
load('vis_g.Rda')
```

The nodes in the network represent researchers with whom I have published. Colors indicate the institution in which I was placed when I first published with each co-author.

|  **UB:** Universitat de Barcelona, Barcelona, Spain
|  **UAB:** Universitat Autonoma de Barcelona,Bellaterra del Valles, Sapin
|  **UofA:** University of Alberta, Edmonton, Canada
|  **UCSF:** University of California San Francisco, San Francisco, USA
<br>

|  **Try:**
|      Zoom in with your mouse scroll function
|      Hover over a node
|      Grab a node and move it around
|      Click on a node to see its connections
<br>

Created in R using [easyPubMed](https://cran.r-project.org/web/packages/easyPubMed/index.html), [tidyverse](https://www.tidyverse.org/), [igraph](https://igraph.org/redirect.html) and [visNetwork](https://datastorm-open.github.io/visNetwork/)


[Learn how to make yours!](/post/network-coauthors/network.html)
<br>
```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
visNetwork(nodes = vis_g$nodes, edges = vis_g$edges, height = "700px", width = '900px')%>%
  visNodes(size = 10, shape='cicle', mass=2,font=list(color='white'),fixed = F) %>%
  visEdges(length = 10,physics = T, smooth = list(enabled=T))%>%
  visLayout(randomSeed = 666)%>%
  visOptions(highlightNearest = list(hover = T), 
             nodesIdSelection = list(enabled = TRUE))%>%
  visLegend()%>%
  visPhysics(stabilization = T, minVelocity = 10, maxVelocity = 25)
```
<br>