---
title: "Bipartite network of authors and research topics"
description: |
  A short description of the post.
author:
  - name: Abel Torres Espin
    url: {}
date: 2024-12-16
categories:
  - Networks
output:
  distill::distill_article:
    self_contained: false
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Libraries
```{r, eval=FALSE}
library(scholar) # For parsing Google scholar information from authors
library(tidyverse)
library(easyPubMed) # For querying PubMed
library(pbapply)
library(parallel)

# dealing with text
library(textclean)
library(tm)
library(SnowballC)

# topic model
library(tidytext)
library(topicmodels)
library(textmineR)

# plotting
library(factoextra)
library(igraph)
library(ggnetwork)
```

```{r, include=FALSE}
library(tidyverse)
library(igraph)
library(ggnetwork)
library(textmineR)
```


```{r, include=FALSE, eval=FALSE}
raw_df<-read.csv("data.csv", na.strings = "") # load csv with data, 3 columns: names, research interest, google id
colnames(raw_df)[1]<-"name" # correct name
```

# Extract authors and abstract information

Starting with the list of papers
```{r, eval=FALSE}
g_id_df<-raw_df%>%
  filter(!is.na(g_scholar))

# API call to get publications ids for each author
pubs<-lapply(g_id_df$g_scholar, get_publications)

# Add names of the author to each df
pubs_name<-lapply(1:length(pubs), function(i){
  pubs[[i]]$name<-g_id_df$name[i]
  pubs[[i]]$id<-g_id_df$g_scholar[i]
  pubs[[i]]
})

# collaps to a single dataset and save in disk
pubs_df<-bind_rows(pubs_name)
saveRDS(pubs_df, file = "pubs_df.Rds")
```

Now we can extract the info per paper. The google schoolar URL connection was a very low time/frequency limit for making requests, so that makes impossible to extract all the information. But we can use the pubmed API.

```{r, eval=FALSE}
pubs_df_f<-pubs_df%>%
  filter(year>=2013)

cl<-makeCluster(14)
parallel::clusterEvalQ(cl, expr = {
  library(easyPubMed)
})
clusterExport(cl, varlist = "pubs_df_f")

pubmed_data<-pblapply(1:nrow(pubs_df_f), function(i){
  pubmed_ids<-get_pubmed_ids_by_fulltitle(pubs_df_f$title[i])
  if(pubmed_ids$Count==0){
    return(NA)
  }else{
    fetch_pubmed_data(pubmed_ids, format = "medline" ) 
  }
}, cl = cl)

saveRDS(pubmed_data, "pubmed_data.Rds")
```

## Process the abstracts
```{r, eval=FALSE}
extract_abs<-function(x){

  if (is.na(x)) return (NA)
  
  start<-which(str_detect(x, "^AB  -"))
  
  if (length(start) == 0){
    return(NA)
  }
  
  count = start + 1
  while( TRUE ){
    if (str_detect(x[count], "^ ")){
      count = count +1
    }else{
      break()
    }
  }
  
  abstract<-paste0(x[start:(count-1)],collapse = "")
  abstract<-str_remove(abstract,"^AB  -")
  
  return(abstract)
}

# abstracts_l<-list()
pubs_df_f$clean_abs<-NA

for (a in 1:length(pubmed_data)){
  cat("\r", a)
  pubs_df_f$clean_abs[a]<-extract_abs(pubmed_data[[a]])
}
```

```{r, eval = FALSE}
# Function for cleaning
clean_text<- function(t){
  t <- as.character(t)
  
  t <- t %>%
    str_to_lower() %>%  # convert all the string to low alphabet
    replace_contraction() %>% # replace contraction to their multi-word forms
    replace_internet_slang() %>% # replace internet slang to normal words
    replace_word_elongation() %>% # replace informal writing with known semantic replacements
    replace_number(remove = T) %>% # remove number
    replace_date(replacement = "") %>% # remove date
    replace_time(replacement = "") %>% # remove time
    str_remove_all(pattern = "[[:punct:]]") %>% # remove punctuation
    str_squish() %>% # reduces repeated whitespace inside a string.
    str_trim() # removes whitespace from start and end of string
  
  return(as.data.frame(t))
}

pubs_df_f$clean_abs<-clean_text(pubs_df_f$clean_abs)$t
pubs_df_f$clean_title<-clean_text(pubs_df_f$title)$t

saveRDS(pubs_df_f, "pubs_df_f.Rds")
```

Adding my abstracts
```{r, eval = FALSE}
ATE_pubs_ids<-get_pubmed_ids("abel torres espin")
ATE_pubmed_data<-fetch_pubmed_data(ATE_pubs_ids, format = "medline")

# get PMIDs
starts<-which(str_detect(ATE_pubmed_data, "^PMID"))
pmid<-ATE_pubmed_data[starts]

#extract papers
starts<-c(starts, length(ATE_pubmed_data))
ATE_pubmed_data_list<-list()
for(i in 1:(length(starts)-1)){
  ATE_pubmed_data_list[[i]]<-ATE_pubmed_data[starts[i]:(starts[i+1]-1)]
}

pubmed_data<-readRDS("pubmed_data.Rds")
pubmed_data<-append(pubmed_data,ATE_pubmed_data_list)

# Save again with the addition of my abstracts
saveRDS(pubmed_data, "pubmed_data.Rds")
```

Adding my information on the pub_df_f
```{r, eval=FALSE}
pubs_df_f<-readRDS("pubs_df_f.Rds")
ATE_pub_df<-data.frame(name = "Abel Torres Espin", 
                       clean_abs = unlist(lapply(ATE_pubmed_data_list, extract_abs)),
                       pubid = pmid)
ATE_pub_df$clean_abs<-clean_text(ATE_pub_df$clean_abs)$t

pubs_df_f<-bind_rows(pubs_df_f, ATE_pub_df)
```

# Text clustering
```{r, eval=FALSE}
pubs_df_clustering<-pubs_df_f%>%
  filter(!str_detect(clean_abs, "this corrects the article")|is.na(clean_abs))

pubs_df_clustering<-pubs_df_clustering%>%
  mutate(clean_abs2 = ifelse(is.na(clean_abs), clean_text(title)$t, clean_abs),
         clean_abs2 = removeWords(clean_abs2, c("health","care", 
                                     "use","studi", "data", "associ", "result",
                                     "research")))
  
pubs_df_clustering<-pubs_df_clustering%>%
  filter(!str_detect(clean_abs2, " des | de | la "))

# create document term matrix (tokens)
set.seed(55)
dtm_clean_abs<-CreateDtm(pubs_df_clustering$clean_abs2,pubs_df_clustering$pubid, 
                    ngram_window = c(1,1), 
                    stopword_vec = stopwords("en"),  
                    verbose = F,
                    stem_lemma_function = function(x) 
                      SnowballC::wordStem(x, "en"))

## Save so that this doesn't need to run every time
saveRDS(dtm_clean_abs, "dtm_clean_abs.Rds")
```

```{r}
dtm_clean_abs<-readRDS("dtm_clean_abs.Rds")

tf_mat <- TermDocFreq(dtm_clean_abs)
## filter for tokens in more than one document
tf_mat_f<-tf_mat%>%
  filter(doc_freq>1 & doc_freq<663)%>%
  filter(!term%in%c("n", "p", "patient", "â", "factor", "increas", "use", "intervent",
                    "studi", "result", ""))

# TF-IDF and cosine similarity
tfidf <- t(dtm_clean_abs[ , tf_mat_f$term ]) * tf_mat_f$idf
tfidf <- t(tfidf)
```

```{r}
csim <- tfidf / sqrt(rowSums(tfidf * tfidf))
csim <- csim %*% t(csim)

cdist <- as.dist(1 - csim)

cdist[is.na(cdist)]<-1

hc <- hclust(cdist, method = "ward.D2")
plot(hc, labels = FALSE, hang = -1)

cdist_mds<-cmdscale(cdist, eig = T, k = 2)
plot(cdist_mds$eig[1:100])
plot(cdist_mds$points, col = cutree(hc, 10))
```

## Explain clusters
```{r}
tfidf_df<-as.data.frame(as.matrix(tfidf))
tfidf_df$clust<-cutree(hc, 60)

tfidf_df_sum<-tfidf_df%>%
  group_by(clust)%>%
  summarise(across(everything(), \(x) mean(x,na.rm = TRUE)))%>%
  pivot_longer(cols = -clust)
  
tfidf_df_sum_f<-tfidf_df_sum%>%
  group_by(clust)%>%
  slice_max(value, n = 5)%>%
  filter(value>0)

tfidf_df_sum_f %>%
  mutate(name = reorder_within(name, value, clust)) %>%
  ggplot(aes(value, name, fill = factor(clust))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ clust, scales = "free") +
  scale_y_reordered()

#write.csv(tfidf_df_sum_f, "tfidf_df_sum_f2.csv")
```

## Graphs
```{r, preview = TRUE}

cluster_names<-read.csv("tfidf_df_sum_f2.csv", na.strings = "")%>%
  mutate(clust_level2 = ifelse(is.na(clust_level2), clust_name, clust_level2))%>%
  group_by(clust)%>%
  summarise(clust_name = unique(clust_name),
            clust_level2 = unique(clust_level2))
  
pubs_df_clustering$clust <- cutree(hc, 60)
pubs_df_clustering2<-pubs_df_clustering%>%
  left_join(cluster_names%>%
              select(clust, clust_level2))%>%
  filter(!is.na(clust_level2))

pubs_df_clustering2$name<-str_trim(pubs_df_clustering2$name)
net_df<-pubs_df_clustering2%>%
  select(name, clust_level2)

g<-graph_from_data_frame(net_df, directed = F)
bipartite.mapping(g)

V(g)$type <- bipartite_mapping(g)$type 
V(g)$size <- degree(g)
g<-simplify(g)

gg_net<-ggnetwork(g, layout = igraph::layout.fruchterman.reingold(g))

gg_net<-gg_net%>%
  mutate(type2 = case_when(
    name == "Abel Torres Espin"~"me",
    type == FALSE~"name",
    type == TRUE~"cluster"
  ),
  size2 = ifelse(type == FALSE, 1, 1.1),
  line_size = ifelse(type2 == "me", 0.15, 0.1))

ggplot(gg_net, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(aes(color = type2), alpha=0.4, 
             curvature = 0.1) +
  geom_nodes(aes(color = type2, shape = type2))+
  geom_nodelabel_repel(aes(label = name, fill = type2), 
                       show.legend = F, force = 0.02)+
  scale_color_manual(values = c("grey","green3", "grey"))+
  scale_fill_manual(values = c("lightpink2", "green3", "steelblue1"))+
  theme_blank()
```




